Modern BTree\\ techniques

### 2020-01-19 btree/presentation.tex
dptree

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-leaf-node/.style={btree-node, fill=greenGood}}
\tikzset{btree-branch-node/.style={btree-node, fill=gray!20}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}
\tikzset{smallbtree-node/.style={btree-node, minimum width=0.4cm, minimum height=0.4cm, rounded corners=0.1cm}}
\tikzset{box/.style={draw, dashed, rounded corners, behind path}}
\tikzset{green-box/.style={box, fill=greenGood}}
\tikzset{blue-box/.style={box, fill=blue!20}}
\tikzset{bloom-filter/.style={
    draw,
    isosceles triangle,
    minimum width=1.2cm,
    shape border rotate=-90,
    isosceles triangle apex angle=60,
    rounded corners
}}
\tikzset{bw-delta/.style={draw, fill=redBad, rounded corners}}
\tikzset{bw-page/.style={draw, fill=greenGood, rounded corners}}

@@

@@

    \begin{overprint}[12cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \node[btree-branch-node] (node1) {};

        \onslide<2>
        \begin{tikzpicture}[]
            \node[btree-branch-node] (node1) {};
            \coordinate[right=0.25cm of node1.center] (link-anchor1);
            \node[below=-0.05cm of link-anchor1] (link1) {\footnotesize{1}};

            \node[btree-branch-node, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-branch-node, below=0.5cm of node1.south] (node3) {};
            \coordinate[right=0.25cm of node3.center] (link-anchor2);
            \node[below=-0.05cm of link-anchor2] (link2) {\footnotesize{2}};

            \node[btree-branch-node, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

@@

    \begin{overprint}[12cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \node[btree-branch-node] (node1) {};

        \onslide<2>
        \begin{tikzpicture}[]
            \node[btree-branch-node] (node1) {};

            \node[btree-leaf-node, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-leaf-node, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};
            \coordinate[above=0.25cm of node5] (left-point);
            \coordinate[above=0.25cm of node10] (right-point);

            \draw[dashed] ([xshift=-2cm]left-point) -- ([xshift=2cm]right-point);
        \end{tikzpicture}

        \onslide<3>
        \begin{tikzpicture}[]
            \node[btree-branch-node] (node1) {};

            \node[btree-leaf-node, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-leaf-node, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};
            \coordinate[above=0.25cm of node5] (left-point);
            \coordinate[above=0.25cm of node10] (right-point);

            \draw[dashed] ([xshift=-2cm]left-point) -- ([xshift=2cm]right-point);

            \smallbtree{1}{1.5}{left=6cm of node1}{}{}{blue-box}
            \node[draw, minimum width=0.5cm, minimum height=0.5cm, rounded corners=0.1cm, below=1.5cm of nodebox1.south] (log2) {};
            \node[draw, minimum width=0.5cm, minimum height=0.5cm, rounded corners=0.1cm, right=0.5cm of log2] (log3) {};
            \node[draw, minimum width=0.5cm, minimum height=0.5cm, rounded corners=0.1cm, left=0.5cm of log2] (log1) {};
            \draw[->, btree-line] (nodebox1.south) -- (log1);
            \draw[->, btree-line] (nodebox1.south) -- (log2);
            \draw[->, btree-line] (nodebox1.south) -- (log3);
        \end{tikzpicture}

    \end{overprint}
    \end{center}
\end{frame}

>>>>

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-leaf-node/.style={btree-node, fill=greenGood}}
\tikzset{btree-branch-node/.style={btree-node, fill=gray!20}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}
\tikzset{smallbtree-node/.style={btree-node, minimum width=0.4cm, minimum height=0.4cm, rounded corners=0.1cm}}
\tikzset{box/.style={draw, dashed, rounded corners, behind path}}
\tikzset{green-box/.style={box, fill=greenGood}}
\tikzset{blue-box/.style={box, fill=blue!20}}
\tikzset{bloom-filter/.style={
    draw,
    isosceles triangle,
    minimum width=1.2cm,
    shape border rotate=-90,
    isosceles triangle apex angle=60,
    rounded corners
}}
\tikzset{bw-delta/.style={draw, fill=redBad, rounded corners}}
\tikzset{bw-page/.style={draw, fill=greenGood, rounded corners}}
\tikzset{dp-tree-log/.style={draw, minimum width=0.5cm, minimum height=0.5cm, rounded corners=0.1cm}}

@@

\newcommand{\basicbtree}[1] {
    \node[btree-branch-node] (node1) {};

    \ifthenelse{ \equal{#1}{show-connections} } {
        \draw[->, btree-line] (node1) -- (node2);
        \draw[->, btree-line] (node1) -- (node3);
        \draw[->, btree-line] (node1) -- (node4);

        \draw[->, btree-line] (node4) -- (node9);
        \draw[->, btree-line] (node4) -- (node10);
    } {}
}

\newcommand{\basicleafconnection}[1] {
    \ifthenelse{ \equal{#1}{active} } {
        \draw[->, btree-line] (node5) -- (node6);
        \draw[->, btree-line] (node6) -- (node7);
        \draw[->, btree-line] (node7) -- (node8);
        \draw[->, btree-line] (node8) -- (node9);
        \draw[->, btree-line] (node9) -- (node10);
    } {
        \draw[->, btree-line, color=gray] (node5) -- (node6);
        \draw[->, btree-line, color=gray] (node6) -- (node7);
        \draw[->, btree-line, color=gray] (node7) -- (node8);
        \draw[->, btree-line, color=gray] (node8) -- (node9);
        \draw[->, btree-line, color=gray] (node9) -- (node10);
    }
}

@@

    \begin{overprint}[12cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
        \end{tikzpicture}

        \onslide<2>
        \begin{tikzpicture}[]
            \basicbtree{}
            \coordinate[right=0.25cm of node1.center] (link-anchor1);
            \node[below=-0.05cm of link-anchor1] (link1) {\footnotesize{1}};

            \coordinate[right=0.25cm of node3.center] (link-anchor2);
            \node[below=-0.05cm of link-anchor2] (link2) {\footnotesize{2}};

@@

    \begin{overprint}[14cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}

            \coordinate[above=0.25cm of node5] (left-point);
            \coordinate[above=0.25cm of node10] (right-point);

            \draw[color=white, opacity=0] ([xshift=-5cm]left-point) -- ([xshift=1cm]right-point);

            \smallbtree{1}{1.5}{left=6cm of node1}{color=white}{color=white}{color=white}

            \coordinate[left=1.2cm of nodebox1.south west] (inserts-start);
            \coordinate[left=1.2cm of nodebox1.north west] (reads-start);
            \node[arrow-pointer, above=0.2cm of inserts-start, color=white] (inserts) {\footnotesize{Insert}};
            \node[arrow-pointer, below=0.2cm of reads-start, color=white] (reads) {\footnotesize{Read}};
        \end{tikzpicture}

        \onslide<2>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}

            \coordinate[above=0.25cm of node5] (left-point);
            \coordinate[above=0.25cm of node10] (right-point);

            \draw[color=white, opacity=0] ([xshift=-5cm]left-point) -- ([xshift=1cm]right-point);

            \smallbtree{1}{1.5}{left=6cm of node1}{}{}{blue-box}

            \coordinate[left=1.2cm of nodebox1.south west] (inserts-start);
            \coordinate[left=1.2cm of nodebox1.north west] (reads-start);
            \node[arrow-pointer, above=0.2cm of inserts-start, color=white] (inserts) {\footnotesize{Insert}};
            \node[arrow-pointer, below=0.2cm of reads-start, color=white] (reads) {\footnotesize{Read}};
        \end{tikzpicture}

        \onslide<3>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
            \basicleafconnection{active}

            \coordinate[above=0.25cm of node5] (left-point);
            \coordinate[above=0.25cm of node10] (right-point);

            \draw[dashed] ([xshift=-5cm]left-point) -- ([xshift=1cm]right-point);

            \smallbtree{1}{1.5}{left=6cm of node1}{}{}{blue-box}

            \coordinate[left=1.2cm of nodebox1.south west] (inserts-start);
            \coordinate[left=1.2cm of nodebox1.north west] (reads-start);
            \node[arrow-pointer, above=0.2cm of inserts-start, color=white] (inserts) {\footnotesize{Insert}};
            \node[arrow-pointer, below=0.2cm of reads-start, color=white] (reads) {\footnotesize{Read}};
        \end{tikzpicture}

        \onslide<4>
        \begin{tikzpicture}[]
            \basicbtree{show-connections}
            \basicleafconnection{}

            \coordinate[above=0.25cm of node5] (left-point);
            \coordinate[above=0.25cm of node10] (right-point);

            \draw[dashed] ([xshift=-5cm]left-point) -- ([xshift=1cm]right-point);

            \smallbtree{1}{1.5}{left=6cm of node1}{}{}{blue-box}
            \node[dp-tree-log, below=1.5cm of nodebox1.south] (log2) {};
            \node[dp-tree-log, right=0.5cm of log2] (log3) {};
            \node[dp-tree-log, left=0.5cm of log2] (log1) {};
            \draw[->, btree-line] (nodebox1.south) -- (log3);
            \draw[->, btree-line] (log1) -- (log2) -- (log3);

            \coordinate[left=1.2cm of nodebox1.south west] (inserts-start);
            \coordinate[left=1.2cm of nodebox1.north west] (reads-start);
            \node[arrow-pointer, above=0.2cm of inserts-start] (inserts) {\footnotesize{Insert}};
            \node[arrow-pointer, below=0.2cm of reads-start] (reads) {\footnotesize{Read}};

            \node[arrow-pointer, left=2cm of node1] {\footnotesize{Read}};
            \node[arrow-pointer, right=0.2cm of nodebox1.south east, rotate=-45, fill=white] {\footnotesize{Merge}};
        \end{tikzpicture}

    \linespread{0.5}
    \vspace{0.5cm}
    \color{black}\fontsize{6pt}{0}\selectfont
        Zhou, Xinjing Shou, Lidan Chen, Ke Hu, Wei Chen, Gang. (2019).  DPTree:
        differential indexing for persistent memory. Proceedings of the VLDB
        Endowment.
    \linespread{1.5}
